Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
names(pf)
qplot(x = dob_day, data = pf ) +
        scale_x_discrete(breaks = 1:31)
# Following three line of code is for different themes
install.packages('ggthemes', dependencies = TRUE)
library(ggthemes) 
theme_set(theme_minimal(24))
# For more theme examples https://github.com/jrnold/ggthemes#examples

```

***

### Faceting
Notes:

```{r Faceting}
qplot(x = dob_day, data = pf ) +
  scale_x_discrete(breaks = 1:31) +
  facet_wrap(~dob_month,ncol = 3) 

```

### Moira's Outlier
Notes:
  - She got and bulk of data in the few hundreds area but some people are just said up to 10 million that was not a big problem for her. She claimed that it is not possible, and throwed that data.
#### Which case do you think applies to Moira’s outlier?
Response: Bad data about an extreme case

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0,1000))
#OR
qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0, 1000, 50))
```

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
# !is.na(gender) means that take the data without na values in the gender.

# Another way to get rid of the NA is omit() function
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
# In here we take the date from pf with omitting the na values.
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure/365, data = pf, binwidth = 1, color = I('black'), fill = I('#F79420'))

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x = tenure, data = pf, binwidth = 30, color = I('black'), fill = I('#099DD9'))

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x = tenure/365, data = pf, xlab = "Number of years using Facebook", ylab = "Number of users in sample", color = I("black"), fill = I("#F79324")) +
  scale_x_continuous(breaks = seq(1,7,1), lim = c(0,7))
```

***

### User Ages
Notes:

```{r User Ages}
# Create a histogram of user ages
qplot(x = age, data = pf, binwidth = 1, 
      color = I('black'), fill = I('#5760a8')) +
  scale_x_discrete(breaks = seq(0,113,5))
# All about age...
summary(pf$age)
```

***

```{r Add a Scaling Layer}
logScale <- qplot(x = log10(friend_count), data = pf)
countScale <- ggplot(aes(x = friend_count), data = pf)+
  geom_histogram() +
  scale_x_log10()
grid.arrange(logScale, countScale, ncol = 2)

```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)),
      xlab = "Friend Count",
      ylab = "Proportion of Users with that friend count",
      binwidth = 10, geom = "freqpoly", color = gender) +
  scale_x_continuous(lim = c(0,1000), breaks = seq(0,1000,50))
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
# Seeing the likes of in tems of gender in Facebook
qplot(x = www_likes, data = subset(pf, !is.na(gender)),
      geom = "freqpoly", color = gender)+
  scale_x_continuous() +
  scale_x_log10() 

# Gives us total likes of men, and women
by(pf$www_likes, pf$gender, sum)
# Gives all statistical info about www_likes in terms of gender
by(pf$www_likes, pf$gender, summary)
```


***

### Box Plots
Notes:

```{r Box Plots}
# Creates a box plot with gender and friend_count varibles
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = "boxplot")
# The little box is our good data
# Other dots are outliers, 
# We will get rid of the outliers to focus on good data
# First way
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = "boxplot", ylim = c(0,1000)) 
# Second way
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = "boxplot") +
  scale_y_continuous(limits = c(0, 1000))
# Third way
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = "boxplot") +
  coord_cartesian(ylim = c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = "boxplot") +
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)

# Who made more friend request?
qplot(x = gender, y = friendships_initiated,
      data = subset(pf, !is.na(gender)), geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 150))
by(pf$friendships_initiated, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
# We should take a closer look to see more detailed

qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = "boxplot") +
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)

# Who made more friend request?
qplot(x = gender, y = friendships_initiated,
      data = subset(pf, !is.na(gender)), geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 150))
by(pf$friendships_initiated, pf$gender, summary)
```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!